Missing The Point:

### 2020-07-15 ICML/MissingThePoint.Rmd
Add pdf version to upload

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "ICML ARTEMISS workshop 2020"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    seal: false
    self_contained: true
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

@@

# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
library(mice)
#load("example_diagnostics.Rdata")
# load("../traceplotdata.Rdata")
# set.seed(12)
# d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
# save(d75, file = "imps.Rdata")
load("imps.Rdata")

# define colorblind friendly colors
paint5 <- c('#228833', '#66CCEE', '#CCBB44','#EE6677', '#AA3377')
# paint2 <- c('#66CCEE', '#EE6677')
```

@@

.right-column[ 
# Algorithmic Convergence

<!-- .footer[Missing The Point | Oberman, van Buuren, Vink | 2020 | github.com/hanneoberman/MissingThePoint] -->

```{r non-conv, echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
p <- gg.mids(d75, x="Y")[[1]] 
p_up <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    x = "Iteration",
    y = "Chain mean", 
    colour = "Imp.")

```{r echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
ggplotly(p_up)
```
]
---

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=$ 1000, $n_{\rm sim}=$ 1000), we obtain valid regression estimates after 5 to 10 iterations. <!-- Multivariate normal set ( $n = 1000$ ), 'amputed' with a varying proportion of incomplete cases, and 'imputed' with a varying number of iterations. Depicted are the bias and coverage rate of an estimated regression coefficient ( $n_{\rm sim} = 1000$ ). -->

```{r results, echo=FALSE, fig.width=9.8, message=FALSE, warning=FALSE}
load("complete.Rdata") 
#library(patchwork)

# create plots
est_bias <- results %>% ggplot(aes(x = t, y = bias.est.X1, color = as.factor(p*100))) +
  geom_hline(yintercept = 0,
             color = "grey",
             lwd = 1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  labs(
    #title = "Average bias in regression estimate", 
    #subtitle = "Multivariate normal data (n = 1000, nsim = 1000)", 
    x = "Number of iterations", 
    y = "Bias (%)", 
    colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "top")

est_cov <- results %>% ggplot(aes(x = t, y = cov.est.X1*100, color = as.factor(p*100))) +
  geom_hline(yintercept = 95,
             color = "grey",
             lwd = 1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  xlab("Number of iterations") +
  ylab("Coverage rate (%)") +
  labs(colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "")

# plot estimates, save 6.75x4.5in or 5.5
# est_bias + est_cov + est_AC + est_Rh + plot_annotation(tag_levels = "A", tag_suffix = ".") + plot_layout(ncol = 2, guides = "collect")

.pull-left[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_bias)
```
]

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_cov)
```
]
]
---

.right-column[ 
# Simulation Study (2)

Whereas non-convergence diagnostics (potential scale reduction factor, auto-correlation) identify signs of non-convergence up-to 30 to 50 iterations.

@@

.pull-right[
<img src="https://github.com/gerkovink/shinyMice/blob/edits/ICML/shinymicehex.png?raw=true" width="255">
]
]
<!-- .footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).] -->

>>>>

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "ICML ARTEMISS workshop 2020"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    seal: false
    self_contained: false
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

@@

# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
library(mice)
#load("example_diagnostics.Rdata")
# load("../traceplotdata.Rdata")
# set.seed(12)
# d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
# save(d75, file = "imps.Rdata")
load("imps.Rdata")
load("complete.Rdata")

# define colorblind friendly colors
paint5 <- c('#228833', '#66CCEE', '#CCBB44','#EE6677', '#AA3377')
# paint2 <- c('#66CCEE', '#EE6677')

# traceplot
p <- gg.mids(d75, x="Y")[[1]] 
p_up <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    x = "Iteration",
    y = "Chain mean", 
    colour = "Imp.")

# create plots
est_bias <- results %>% ggplot(aes(x = t, y = bias.est.X1, color = as.factor(p*100))) +
  geom_hline(yintercept = 0,
             color = "grey",
             lwd = 1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  labs(
    #title = "Average bias in regression estimate", 
    #subtitle = "Multivariate normal data (n = 1000, nsim = 1000)", 
    x = "Number of iterations", 
    y = "Bias (%)", 
    colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

est_cov <- results %>% ggplot(aes(x = t, y = cov.est.X1*100, color = as.factor(p*100))) +
  geom_hline(yintercept = 95,
             color = "grey",
             lwd = 1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  xlab("Number of iterations") +
  ylab("Coverage rate (%)") +
  labs(colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

.left-column[ 
```{r echo=TRUE, eval=FALSE}
Missing The Point

.right-column[

# Algorithmic Convergence

```{r echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
ggplotly(p_up)
```
]

---

```
]

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=$ 1000, $n_{\rm sim}=$ 1000), we obtain valid regression estimates after 5 to 10 iterations.

.pull-left[
```{r echo=FALSE, fig.height=5.4, fig.width=5.1, message=FALSE, warning=FALSE}
ggplotly(est_bias)
```
]

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_cov)
```
]
]

---

@@

.right-column[

# Simulation Study (2)

Whereas non-convergence diagnostics (potential scale reduction factor and auto-correlation) identify signs of non-convergence up-to 30 to 50 iterations.

@@

.pull-right[
<img src="https://raw.githubusercontent.com/gerkovink/shinyMice/edits/shinymicehex.png" width="255">
]
]
<!-- .footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).] -->
